---
title: "Intro to Tidyverse Workshop"
author: "Bin Chen"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
    theme: cosmo
  word_document:
    toc: yes
---
Welcome to the "Introduction to Tidyverse" workshop! This is the second training of the two sequence R workshop this fall. In this training, we will (1) review key concepts from the first training (e.g., working directory, data types and structures); (2) use most common commands about the R package `tidyverse`; and (3) walk through a standard data analysis flow by using the `tidyverse` package and knit `.RMD` file to a `HTML` file.  

For working directory issues, please follow the instructions to set your Rstudio:

- Click "Tools" in Rstudio, and click "Global Options" from the option lists. 
- Click "R Markdown" from the pop-up window "Options". 
- Look for "Evaluate chunks in directory"   
- Change the default option "Document" to "Project"

Then do the followings: 
- Create a new R project, and three folders `data`, `R`, and `output`. 
- Place the input data file `gamesale.csv` in the `data` subfolder.
- Place the RMD file `tidyverse_workshop_incomplete.csv` in the `R` subfolder.

The dataset `gamesale.csv` includes the world-wide sales data of video games. The description of the variables included in the data file is as follows: 

- `Rank` - Ranking of overall sales
- `Name` - Video game names
- `Platform` - Platform of the games release (i.e. PC,PS4, etc.)
- `Year` - Year of the game's release
- `dual_sim` - Has dual sim support or not
- `Genre` - Genre of the game
- `Publisher` - Publisher of the game
- `NA_Sales` - Sales in North America(in millions)
- `EU_Sales` - Sales in Europe(in millions)
- `JP_Sales` - Sales in Japan(in millions)
- `Other_Sales` - Sales in other areas of the world(in millions)
- `Global_Sales` - Total worldwide sales(in millions)

For more information, you could refer to this webpage [https://www.kaggle.com/gregorut/videogamesales], from which the dataset was downloaded. 

Credit to Dr. Seung Choi for introducing this dataset. 


# What is `tidyverse`?

The tidyverse is a collection of R packages designed for data science.

A list of packages that we used most but not all

*   `readr`
*   `dplyr`
*   `tidyr`
*   `stringr`
*   `ggplot2`

For more information, visit [https://www.tidyverse.org/packages/]

# R Data Analysis Workflow 

## Import Data

*   base R or `readr`: for flat files (e.g., `.csv`)
*   `jsonlite` package for JSON data
*   `haven` package for SPSS, STATA, and SAS data

## Data Wrangling 

*   `tidyr`
*   `dplyr`
*   `lubridate` for dates and date-times data

## Data Visualization 

Normally, you will want to do some statistical modeling or other analysis after data cleaning and wrangling. But we will not discuss it here.

*   `ggplot2`


# Preparation

## Install the package 

Run `install.packages("tidyverse")` in console. Because you only need to install it for once.
Run `library(tidyverse)` to load the package

```{r}
library(tidyverse)
```

## Forward pipe operator `%>%`

The infix operator `%>%` is not part of base R, but is in fact defined by the package `magrittr` and is heavily used by `dplyr`.

```{r}
x <- c("high", "low", "high")
```

```{r}
x1 <- as.factor(x)
```

```{r}
x2 <- x %>%
  as.factor() 
```

```{r}
x1 == x2 
```

# Commands

You may need to use hundreds of commands in a single task or when dealing with different types of data. Here we will only talk about the mostly used ones in standard scenarios. 

## Import data `read_csv`

```{r}
df <- read_csv("data/gamesale.csv", na = "N/A")
```

base R command: `read.csv`

```{r}
df0 <- read.csv("data/gamesale.csv")
```

You may not see any differences between these two but the differences exist. One point is `read_csv` is significantly faster for large .csv files. 

There are many options to specify when import data into R (e.g., missing value)

## Check Missing Value

Check NA in the whole data frame use `sum()` and `is.na()`

```{r}
sum(is.na(df))
```

Check NA by column 
```{r}
sum(is.na(df$Year))
```

## Convert data type 

* `as.numeric()`
* `as.factor()`
* `as.character()`

What is the data type of the column `Year`

```{r}
class(df$Year)
```

We want to convert it to a `integer`

```{r}
df$Year <- as.integer(df$Year)
```

```{r}
class(df$Year)
```
Note: Although we did not do it here - it is always a good habit to create a new column whenever you want to manipulate original data.

## Create a new column `mutate`

Lets convert `Platform`, `Genre`, and `Publisher` to `factor` data and create three new columns named `Platform_f`, `Genre_f`, and `Publisher_f`.

hint: use `as.factor`

```{r}
df1 <- df %>%
  mutate("Platform_f" = as.factor(Platform),
         "Genre_f" = as.factor(Genre),
         "Publisher_f" = as.factor(Publisher)
         )
```

Lets check
```{r}
class(df1$Genre_f)
```

Note. You can use `across()` and `is.charater` to do this more efficiently. 

```{r}
df<- df %>%
  mutate(across(where(is.character), as.factor))
```


## Select columns `select()`

Suppose we are only interested in three columns: `Rank`, `Name`, and `Global_sales`

```{r}
df2 <- df %>%
  select (Rank, Name, Global_Sales)
```

Lets check 
```{r}
colnames(df2)
```

## Subset rows with `subset()`

To subset certain rows, you need specify rules. 

Suppose we want to subset top 50 most popular games. 

```{r}
df3 <- df %>%
  subset (Rank <= 50)
```

What if we have multiple conditions? (use `&` and "|")

e.g., Top 50 games on PC 

```{r}
df4 <- df %>%
  subset(Rank <= 50 & Platform == "PC")
```

```{r}
levels(df$Platform)
```


What if we want subset `Platform` is "PC", "PS3", "Wii", "X360"?

Use `%in%` to identify if an element (e.g., a number) belongs to a vector or dataframe

```{r}
df5 <- df %>%
  subset(Platform %in% c("PC", "PS3", "Wii", "X360"))
```

```{r}
unique(df5$Platform)
```


### Subset rows with `filter`

Note. `subset()` and `filter()` do same things but `filter()` is faster when data size is large. For more discussions visit [https://stackoverflow.com/questions/39882463/difference-between-subset-and-filter-from-dplyr]


## `group_by`

Use `group_by` and `count ()`to list number of games for each `Genre`

```{r}
df %>%
  group_by(Genre) %>%
  count()
```

Suppose we want to Compare mean value of `Global_sales` by `Genre`
hint: `summarise`

```{r}
df %>%
  group_by(Genre) %>%
  summarise(sale = mean(Global_Sales))
```

Lets reorder the table above by `sale` from high to low 

```{r}
df %>%
  group_by(Genre) %>%
  summarise(sale = mean(Global_Sales)) %>%
  arrange(-sale) # notice the negative sign here
```

## Visualization `ggplot2`

Suppose we want to plot the global sales by game genres

`ggplot()`
`aes()`
`geom_bar`: `stat=`

Note. To add more layers to a plot, you use `+` rather than `%>%`

```{r}
df %>%
  ggplot(aes(x = Genre, y = Global_Sales)) + # note "+" not "%>%"
  geom_bar(stat='identity') # 'identity' means y value is y = global_sales
```

Lets make this plot more informational and prettier by adding title, axis lable, and set a theme style

```{r}
df %>%
  ggplot(aes(x = Genre, y = Global_Sales)) + 
  geom_bar(stat='identity') +
  ggtitle("Video Game Global Sales by Genre") + # title
  ylab ("Global Sale (Million)") + # ylab 
  theme_minimal() # theme
```

# Gamesale data showcase

## Load packages

Besides of `tidyverse`, we will use two more packages here

`skimr`: provides a strong set of summary statistics.
`kableExtra`: generates tables (upgrade base R `kable`)

You should see the warning upon opening this RMD file of installing packages.
If you didn't do it, use `install.package()` to install packages.

```{r message = FALSE}
library(kableExtra) # use for table
```

## Generate a table 
```{r}
df %>%
  subset(Platform %in% c("PS3", "X360", "Wii", "PC")) %>%
  group_by(Platform) %>%
  summarise(mean_sale_Global = mean(Global_Sales),
            sd_sale_Global = sd(Global_Sales),
            max_sale_Global = max(Global_Sales),
            min_sale_Global = min(Global_Sales)) %>%
  arrange(-mean_sale_Global) %>%
  kable(caption = "Table of Sales of Video Games by Platforms") %>%
  kable_classic(full_width = FALSE, html_font = "Arial")
```

## Generate a Plot

Note. You can use `unique()` to check unique values of a column
```{r}
unique(df$Genre)
```

```{r}
df %>%
  filter(Platform %in% c("PS4", "X360", "Wii", "PC")) %>%
  filter(Genre %in% c("Sports", "Role-Playing", "Shooter", "Action")) %>%
  ggplot(aes(fill=Platform, y=Global_Sales, x=Platform)) + 
  geom_bar(position="dodge", stat="identity") +
  theme_minimal() +
  facet_wrap(~Genre) + # present multiple plots
  theme(legend.position="none") +
  xlab("") +
  ylab("Global Sales (million)")
```








